High-Fidelity Counterfactual Explanations through Conformal Prediction

### 2023-04-21 notebooks/synthetic.qmd
ECCCo

        # Classifier and coverage:
        conf_model = conformal_model(clf; method=:simple_inductive, coverage=cov)
        mach = machine(conf_model, X, y)
        fit!(mach)
        M = CCE.ConformalModel(mach.model, mach.fitresult)

        # Set up CCE:
        factual_label = predict_label(M, data, x)[1]
        target_label = data.y_levels[data.y_levels .!= factual_label][1]

@@

            # CCE for given classifier, coverage, temperature and λ:
            generator = CCEGenerator(temp=temp, λ=[l2_λ,λ], opt=opt)
            @assert predict_label(M, data, x) != target_label
            ce = try
                generate_counterfactual(
                    x, target_label, data, M, generator;
                    initialization=:identity,
                    converge_when=:generator_conditions,
                    gradient_tol=gradient_tol,
                    max_iter=max_iter,
                )
            catch
                missing
            end

@@

# Untrained Models:
models = Dict(Symbol("cov$(Int(100*cov))") => CCE.ConformalModel(conformal_model(mlp; method=:simple_inductive, coverage=cov)) for cov in cvgs)

# Measures:
measures = [
    CounterfactualExplanations.distance,
    CCE.distance_from_energy,
    CCE.distance_from_targets,
    CounterfactualExplanations.validity,
]
```

@@

        # Model training:
        M = train(M, data)
        # Set up CCE:
        factual_label = predict_label(M, data, x)[1]
        target_label = data.y_levels[data.y_levels .!= factual_label][1]

@@

            # Generators:
            _generators = deepcopy(generators)
            _generators[:cce] = CCEGenerator(temp=_temp, λ=[l2_λ,λ], opt=opt)
            _generators[:energy] = CCE.EnergyDrivenGenerator(λ=[l2_λ,λ], opt=opt)
            _generators[:target] = CCE.TargetDrivenGenerator(λ=[l2_λ,λ], opt=opt)

@@

                # CCE for given models, λ and generator:
                @assert predict_label(M, data, x) != target_label
                ce = try
                    generate_counterfactual(
                        x, target_label, data, M, gen;
                        initialization=:identity,
                        converge_when=:generator_conditions,
                        gradient_tol=gradient_tol,
                        max_iter=max_iter,
                    )
                catch
                    missing
                end

@@

```{julia}
bmks = []
for (dataname, dataset) in datasets
    for λ in Λ, temp in temps
        _generators = deepcopy(generators)
        _generators[:cce] = CCEGenerator(temp=temp, λ=[l2_λ,λ], opt=opt)
        _generators[:energy] = CCE.EnergyDrivenGenerator(λ=[l2_λ,λ], opt=opt)
        _generators[:target] = CCE.TargetDrivenGenerator(λ=[l2_λ,λ], opt=opt)
        bmk = benchmark(
            dataset; 
            models=deepcopy(models), 
            generators=_generators, 
            measure=measures,
            suppress_training=false, dataname=dataname,
            n_individuals=5,
            initialization=:identity,
        )
        bmk.evaluation.λ .= λ
        bmk.evaluation.temperature .= temp
        push!(bmks, bmk)
    end
end
bmk = reduce(vcat, bmks)
```

>>>>

        # Classifier and coverage:
        conf_model = conformal_model(clf; method=:simple_inductive, coverage=cov)
        mach = machine(conf_model, X, y)
        fit!(mach)
        M = ECCCo.ConformalModel(mach.model, mach.fitresult)

        # Set up ECCCo:
        factual_label = predict_label(M, data, x)[1]
        target_label = data.y_levels[data.y_levels .!= factual_label][1]

@@

            # ECCCo for given classifier, coverage, temperature and λ:
            generator = ECCCoGenerator(temp=temp, λ=[l2_λ,λ], opt=opt)
            @assert predict_label(M, data, x) != target_label
            ce = try
                generate_counterfactual(
                    x, target_label, data, M, generator;
                    initialization=:identity,
                    converge_when=:generator_conditions,
                    gradient_tol=gradient_tol,
                    max_iter=max_iter,
                )
            catch
                missing
            end

@@

# Untrained Models:
models = Dict(Symbol("cov$(Int(100*cov))") => ECCCo.ConformalModel(conformal_model(mlp; method=:simple_inductive, coverage=cov)) for cov in cvgs)

# Measures:
measures = [
    CounterfactualExplanations.distance,
    ECCCo.distance_from_energy,
    ECCCo.distance_from_targets,
    CounterfactualExplanations.validity,
]
```

@@

        # Model training:
        M = train(M, data)
        # Set up ECCCo:
        factual_label = predict_label(M, data, x)[1]
        target_label = data.y_levels[data.y_levels .!= factual_label][1]

@@

            # Generators:
            _generators = deepcopy(generators)
            _generators[:cce] = ECCCoGenerator(temp=_temp, λ=[l2_λ,λ], opt=opt)
            _generators[:energy] = ECCCo.EnergyDrivenGenerator(λ=[l2_λ,λ], opt=opt)
            _generators[:target] = ECCCo.TargetDrivenGenerator(λ=[l2_λ,λ], opt=opt)

@@

                # ECCCo for given models, λ and generator:
                @assert predict_label(M, data, x) != target_label
                ce = try
                    generate_counterfactual(
                        x, target_label, data, M, gen;
                        initialization=:identity,
                        converge_when=:generator_conditions,
                        gradient_tol=gradient_tol,
                        max_iter=max_iter,
                    )
                catch
                    missing
                end

@@

```{julia}
bmks = []
for (dataname, dataset) in datasets
    for λ in Λ, temp in temps
        _generators = deepcopy(generators)
        _generators[:cce] = ECCCoGenerator(temp=temp, λ=[l2_λ,λ], opt=opt)
        _generators[:energy] = ECCCo.EnergyDrivenGenerator(λ=[l2_λ,λ], opt=opt)
        _generators[:target] = ECCCo.TargetDrivenGenerator(λ=[l2_λ,λ], opt=opt)
        bmk = benchmark(
            dataset; 
            models=deepcopy(models), 
            generators=_generators, 
            measure=measures,
            suppress_training=false, dataname=dataname,
            n_individuals=5,
            initialization=:identity,
        )
        bmk.evaluation.λ .= λ
        bmk.evaluation.temperature .= temp
        push!(bmks, bmk)
    end
end
bmk = reduce(vcat, bmks)
```
